---
title: "EDA_credit_card_marketing_campaign"
author: "Group 18"
date: "9/20/2019"
output: html_document
---

## PART 2 - Extreme Gradient Boosted Trees with XGBoost

#### Load required libraries
```{r, warning=FALSE, message=FALSE}
library(caret)
library(MASS)
library(tidyverse)
library(tidyr)
library(ggplot2)
library(ggthemes)
library(dplyr)
library(mlbench)
library(class)
library(pROC)
library(ROSE)
library(e1071)
library(gplots)
library(glmnet)
library(ISLR)
library(viridis)
library(gridExtra)
library(grid)
library(DMwR)
library(xgboost)
```

### Load Credit Card Dataset

```{r, fig.width=15, fig.height=6, warning=FALSE}
# Loading the data
rm(list=ls())
credit.data.csv <- read.csv("creditcardmarketing-bbm.csv")

# Remove the ID column "Customer.Number" which is not helpful in the model
credit.data.csv <- credit.data.csv[, -1]

# Overview the data
des.data <- matrix(nrow = ncol(credit.data.csv),
                 ncol = 13,
                 dimnames = list(colnames(credit.data.csv),
                                 c("Type", "Records", "MissingValue",
                                   "Min", "1st Quatile", "Median",
                                   "3rd Quatile", "Max", "Mean",
                                   "StardardDeviation", "Skewness", "Kurtosis",
                                   "Description")))

for (i in 1:ncol(credit.data.csv)) {
    des.data[i, 1] <-class(credit.data.csv[, i])
    des.data[i, 2] <-length(unique(credit.data.csv[, i]))
    des.data[i, 3] <-sum(is.na(credit.data.csv[, i]))
}

```

#### Removing missing values

```{r, fig.width=15, fig.height=6, warning=FALSE}
# removing rows with NA values
credit.data <- na.omit(credit.data.csv)
# checking again after removal
which(is.na(credit.data))

# the vector of column numbers for factor-type variables
rnum_factor <- as.vector(which(des.data[,"Type"] == "factor"))
```

### Feature Selection with imbalanced Data (Same as in Part-1)

```{r, Feature engineering, warning=FALSE, message=FALSE}
# imbalance data
table(credit.data$Offer.Accepted)
# due to linear relationship between Average.Balance and quarterly balance
# we will remove the quarterly number to avoid multicollinearity
my.df <- credit.data[,c(1:12)]
for (n in 1:7) {
  my.df[, rnum_factor[n]] <- my.df[,rnum_factor[n]] %>% as.numeric()
}

# Standardize predictor
# 0 - Not accepted
# 1 - Accepted
my.df$Offer.Accepted <- my.df$Offer.Accepted - 1

# dividing househols size in single/family
# 1 = single
# 2 = couple of family = 2 or more
my.df$Family.type[my.df$Household.Size <= 1] <- 1
my.df$Family.type[my.df$Household.Size > 1] <- 2

# create a new feature from Average.Balance divided by Household.Size 
# to measure the financial soundness of the famility
my.df$Average.Balance.Household.Size <- my.df$Average.Balance / my.df$Household.Size

# remove Average.Balance and Household.Size after new feature is being keyed
my.df <- my.df[,c(2:9,11,13,14,1)]
colnames(my.df)
```

### Experiment Setup   

##### Why Under Sampling and XGBoost ?      
- Looking at part 1 results for different models with various sampling method, we conclude that under-sampling gace "balanced" results between specificy and sensitivity.
- We decided to further explore under sampling with one of the most popular models at the moment i.e. XGBoost.
- XGBoost is one of top implementations of Gradient Boosting and the weighted approach might help us to improve results especially with improvement in Sensitivity (True Positive Rate) and Specificity (True Negative Rate)

##### Train/Test split   
- As under-sampling gives us limited training data, we tried train-test split of 80-20 and 75-25.
- Based on evaluation metric results, 80-20 performed better and hence we prefer to use 80-20 split approach on this dataset.

##### Feature Selection   
- Feature selection was performed using simple linear model that minimised error
- Top 8 features were selected for pre-processing

```{r, Train Test Split, warning=FALSE, message=FALSE}
set.seed(123)
## split data into 75% train and 25% test
## split data into 80% train and 20% test
inTrain<- createDataPartition(my.df$Offer.Accepted, p= 0.8)[[1]]
mytrain<- my.df[inTrain, ]
mytest<- my.df[-inTrain, ]

# Under Sampling
mytrain.Bal <- ovun.sample(Offer.Accepted ~ ., data = mytrain, method = "under",
                           seed = 123)$data
mytest.Bal <- ovun.sample(Offer.Accepted ~ ., data = mytest, method = "under",
                           seed = 123)$data

# Total positive samples in train data
table(mytrain$Offer.Accepted)
# Check the imbalance of data now
table(mytrain.Bal$Offer.Accepted)
table(mytest.Bal$Offer.Accepted)
cls.mytrain <- mytrain.Bal$Offer.Accepted
cls.mytest<- mytest.Bal$Offer.Accepted

cls.mytrain.Bal <- mytrain.Bal$Offer.Accepted

# Create feature selection pocket
my.features <- c()

# Run Accuracy check for the 1st feature
my.acc.first<-c()
for (i in 1:(ncol(mytrain.Bal) - 1)) {
  model <- lm(data=mytrain.Bal[, c(i, ncol(mytrain.Bal))], Offer.Accepted~.)
  yhat <- ifelse(predict.lm(model, mytest.Bal[, c(i, ncol(mytrain.Bal))]) > 0.5, 1, 0)
  acm <- confusionMatrix(as.factor(yhat),as.factor(cls.mytest))
  a <- acm$overall["Accuracy"]
  my.acc.first<-c(my.acc.first, a)
}
# find the name of the feature who produced minimal MSE
my.acc.first.max <- colnames(mytest.Bal)[which(my.acc.first== max(my.acc.first))][1]
my.features<- c(my.features, my.acc.first.max)
print(my.features)

# Find the next 7 features
my.acc <- my.acc.first
for (j in 1:6) {
  current.max.acc <- -Inf
  selected.i <- NULL

  for(i in 1: (ncol(mytrain.Bal) - 1)) {

    current.f <- colnames(mytrain.Bal)[i]
   
    if(current.f %in% c(my.features, "Offer.Accepted")) {next}
    model <- lm(data=mytrain.Bal[, c(my.features, current.f, "Offer.Accepted")], Offer.Accepted~.)
   
    yhat <- ifelse(predict.lm(model, 
                              mytest.Bal[, c(my.features, current.f, "Offer.Accepted")]) > 0.5, 
                              1, 0)
    # calculate the confusion matrix
    newacm <- confusionMatrix(as.factor(yhat),as.factor(cls.mytest))
    newa <- newacm$overall["Accuracy"]
    # find the max accuracy 
      if(newa > current.max.acc) {
        current.max.acc <- newa
        selected.i <- colnames(mytrain.Bal)[i]
      }
    }
  # append the feature into the vector when max accuracy is found
  my.features <- c(my.features, selected.i)
 
}
# The top 8 features
```

> Top 7 Features are : `r my.features`

### Feature pre-processing
- The top 7 features are categorical variables
- One hot encoding was performed using Keras Dummtvars
- Train and Test data was merged to make sure all levels of samples were represented in encoded set
- A dummy column was created to identify train and test data in combined dataset.
Train=1 -> Training Data
Train=0 -> Test Data

```{r, Train Test, warning=FALSE, message=FALSE}
# https://medium.com/@vaibhavshukla182/how-to-solve-mismatch-in-train-and-test-set-after-categorical-encoding-8320ed03552f
mytrain.Bal <- mytrain.Bal[,c(which(colnames(mytrain.Bal) %in% my.features),12)]
mytest.Bal <- mytest.Bal[,c(which(colnames(mytest.Bal) %in% my.features),12)]

# Create Train / Test Identifier
mytrain.Bal['train'] <- 1
mytest.Bal['train'] <- 0

# Combine Data
combined.data <- rbind(mytrain.Bal, mytest.Bal)
combined.data[,!(names(combined.data) %in% c("Average.Balance.Household.Size",
                                             "Offer.Accepted" , "train"))] <-lapply(combined.data[,!(names(combined.data) %in% c("Average.Balance.Household.Size","Offer.Accepted" ,"train"))], factor)

dummies <- dummyVars(~ ., data =  combined.data[,!(names(combined.data) %in% c("Average.Balance.Household.Size","Offer.Accepted" ,"train"))])

trsf.data <- data.frame(predict(dummies, newdata = combined.data))
trsf.df <- cbind(trsf.data, combined.data[,(names(combined.data) %in% c("Average.Balance.Household.Size","Offer.Accepted" ,"train"))])

# Split data back to train and test
train.data <- trsf.df[trsf.df$train == 1,!(names(trsf.df) %in% c("train"))]
test.data <- trsf.df[trsf.df$train == 0,!(names(trsf.df) %in% c("train"))]
cls.train <- trsf.df$Offer.Accepted[trsf.df$train == 1]
cls.test<- trsf.df$Offer.Accepted[trsf.df$train == 0]
```

## Classification Model

### Modeling Approach   
1. Top 7 features selected from available feature set
2. Data pre-processing with one hot encoding of categorical features
3. Hyper tuning XGBoost with 10-fold and 5-fold cross validation. 
Based on test results, 10-fold cross validation was selected for final modeling.

##### Let's start with HyperParameter Tuning...

```{r, Hyper Parameter Tuning}
set.seed(123)
# Create Search Grid
searchGrid <- expand.grid(eta_values = c(0.01, 0.05, 0.1),
                          max_depth_values = c(4, 6, 8),
                          gamma_values = c(1, 5),
                          min_child_weight_values = c(1, 2,4))

# Test Grid 
# searchGrid <- expand.grid(eta_values = c(0.001),
#                           max_depth_values = c(2,3),
#                           gamma_values = c(5),
#                           min_child_weight_values = c(2))

# xgb Matrix for input
dtrain <- xgb.DMatrix(data = as.matrix(train.data[,!(names(train.data) %in% c("Offer.Accepted"))]),
                              label = train.data$Offer.Accepted)
dtest <- xgb.DMatrix(data = as.matrix( test.data[,!(names(test.data) %in% c("Offer.Accepted"))]),
                             label = test.data$Offer.Accepted)

xgbst.results <- apply(searchGrid, 1, function(parameterList){
        #Extract Parameters to validate
        current_eta <- parameterList[["eta_values"]]
        current_depth <- parameterList[["max_depth_values"]]
        current_gamma <- parameterList[["gamma_values"]]
        current_child_weight <- parameterList[["min_child_weight_values"]]
        
        param <- list(objective = "binary:logistic",
                      eta = current_eta,
                      max_depth = current_depth,
                      gamma = current_gamma,
                      min_child_weight = current_child_weight)
        
        xgb.cv.model <- xgb.cv(params = param,
                               data = dtrain,
                               metrics = "logloss", #minimise logloss
                               nrounds = 500,
                               nfold = 10, # 10 fold gives better result than 5 fold
                               early_stopping_rounds = 15,
                               verbose = F)
        
        return(list(p = param,
                    loss = min(xgb.cv.model$evaluation_log$test_logloss_mean),
                    round = xgb.cv.model$best_iteration))
})
```

### Find Best Parameters

```{r, best params}
best.logloss <- 100
best.params <- c()
best.iter <- 0
for (i in 1:length(xgbst.results)){
      if (xgbst.results[[i]]$loss < best.logloss) {
        best.logloss <- xgbst.results[[i]]$loss
        best.params <- xgbst.results[[i]]$p
        best.iter <- xgbst.results[[i]]$round
      } else {
      }
}

# Best parameters
best.params
# Best iterations for boosting
best.iter
# Minimum loss value
best.logloss
```

### Train model on best parameters

```{r, train and test xgboost}
xgb.model <- xgb.train(params = best.params,
                       data = dtrain,
                       metrics = "logloss",
                       nrounds = best.iter,
                       print_every_n=10,
                       verbose = T)
```

### Evaluation and Conclusion  

##### Pros   
- Based on evaluation metric, Decision Tree and XGBoost give better results with Under Sampling

##### Cons
- All models suffer from low specificity – given trials on various sampling techniques, this points to data incompleteness

##### Conclusion   
- A successful credit card campaign will rely on following parameters
1) How the client is approached? i.e. by a Letter or a Postcard
2) What kind of rewards are on offer? i.e. Air Miles, Cash Back or Points
3) What is the Credit Rating of the client and How many Credit Cards a client holds?
4) Home owenership plays crucial role in client's decision to accept or reject the offer.

##### Further Improvements     
- A relatively low sensitivity scores points to bias prediction value of 1.
The model can be further improved with gathering more successful cases, which can allow us to use under sampling in more effective manner.

```{r, train and test}
offer.pred.prob <- predict(xgb.model, newdata = dtest)
offer.pred <- ifelse (offer.pred.prob > 0.5,1,0)
conf.mat <- confusionMatrix(as.factor(offer.pred),
                      as.factor(test.data$Offer.Accepted))

# Confusion Matrix
conf.mat

x.axis <- c("Accuracy", "F1", "Precision", "Recall", "Sensitivity", "Specificity")
y.axis <- c(conf.mat$overall["Accuracy"],
            conf.mat$byClass["F1"],
            conf.mat$byClass["Precision"],
            conf.mat$byClass["Recall"],
            conf.mat$byClass["Sensitivity"],
            conf.mat$byClass["Specificity"])

d <- data.frame(x = x.axis, y = y.axis)
ggplot(d) +geom_point(aes(x, y), color = "blue") +
  ylim(0,1) +
  xlab(label = "Measurement") +
  ylab(label ="Value") +
  labs(title = "Hypertuned XGBoost with Under Sampling",
       caption = "Group 18, STAT 5003, University of Sydney")
```

### ROC Curve - Sensitivity vs Specificity   

```{r, roc curve}
pROC_obj <- roc(test.data$Offer.Accepted,
                offer.pred,
                smoothed = TRUE,
                # arguments for ci
                ci=TRUE, ci.alpha=0.9, stratified=FALSE,
                # arguments for plot
                plot=TRUE, auc.polygon=TRUE, max.auc.polygon=TRUE, grid=TRUE,
                print.auc=TRUE, show.thres=TRUE)


sens.ci <- ci.se(pROC_obj)
plot(sens.ci, type="shape", col="lightblue")
plot(sens.ci, type="bars")
```

#### Session Information

```{r, session information display}
# sessionInfo()
```
